Make any stupid plot you want

### 2021-01-11 draft/sections/implementation.tex
need to resort fiber section

\subsection{Data: $\mathcal{E}$}
We propose a semantic markup of the fiber and connectivity that we will later use to check the constraints of the $\nu$ and $Q$. 
\begin{minted}{python}
class FiberBundle:
    def __init__(self, K, F):
            """
            K: {'tables':[]}
            F: {'field name': (type, monoid action)}
            """

    def is_section(self, section):
        """checks if a section is from a given fiber bundle:
        are values in F, are keys in K"""
        pass
\end{minted}
We add a monoid field to the schema like structure of the fiber since the monoid actions are the constraints on the $\nu$ and $Q$. We implement $K$ using the triangulization scheme described in section~\ref{sec:triangulization}. This means we expect data to be provided as tables where the name encodes the connectivity:

>>>>

\subsection{Data: $\mathcal{E}$}
We propose a semantic markup of the fiber and connectivity that we will later use to check the constraints of the $\nu$ and $Q$. 
\begin{minted}{python}
class FiberBundle:
    def __init__(self, K, F):
            """
            K: {'tables':[]}
            F: {'field name': {'type': data type, 'monoid': measurement type, 
                                'range': restrictions on data type }}
            """

    def is_section(self, section):
        """checks if a section is from a given fiber bundle:
        are values in F, are keys in K"""
        pass
\end{minted}
We add a monoid field to the schema like structure of the fiber since the monoid actions are the constraints on the $\nu$ and $Q$. These will usually be identified by a measurment scale such as 'nominal', 'ordinal', 'interval', and 'ratio'. The optional? range field can be used to specify restrictions on the type such as set of string categorical values or fixed interval of float values.  We implement $K$ using the triangleization scheme described in section~\ref{sec:triangulization}. This means we expect data to be provided as tables where the name encodes the connectivity:

### 2021-01-11 draft/sections/math.tex
refactored vector simplex + point w/ new fiber structure
need to resort fiber section

%% convention is base space \times fiber bundle
%% change B to K
\section{Notation \& Definitions}
%% intro placeholder of eq in intro
%% use let or where before you introduce symbols
%% more notations w/ general words 
%% tone down - mathcal A, mathcal E, and mathcal H - space, will formalized in this section 
%% give the why of the math after every mention of why
In this section we introduce a mathematical description of the visualization pipeline where artist $A$ functions transform data of type $\Gamma(E)$ to an intermediate representation in prerendered display space of type $\Gamma(H)$:

\begin{equation}
    \label{eq:artist}
    A: \Gamma(E) \rightarrow \Gamma(H)
\end{equation}

\begin{itemize}
\item $A$ is the function that converts an instance of data $\Gamma(E)$ to an instance of a visual representation $\Gamma(H)$ 
\item $E$ is a locally trivial fiber bundle over $K$ representing data space.
\item $H$ is a fiber bundle over $S$ representing visual space
\item $K$ and $S$ is a triangulizable topological space or CW complex encoding the connectivity of the points in $E$
\end{itemize}
All the fiber bundles mentioned in this work is assumed to be locally trivial \cite{rowlandFiberBundle,FiberBundle2020}.

\subsection{Data Fiber Bundles}
\begin{figure}[H]
    \label{fig:data_fiber_bundle}
    \includegraphics[width=.2\linewidth]{figures/sections/math/fiberbundle.png}
    \caption{write up some words here}
\end{figure}

As proposed by Butler \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, we model data as a fiber bundle (E, K, $\pi$, F) with $\pi: E \rightarrow K$ where $K$ which can be thought of as a set of keys $k$. A section $\tau:K \rightarrow E$ is an instance of the data that lies in $E$ and is discussed in section~\ref{sec:data_fiber}.

We provide a brief description of fiber bundles because we model data, visual transformations, and a prerendered visual graphic as fiber bundles.  Fiber bundles are product spaces of toplological spaces, which are a set of points with a set of neighborhoods for each point\cite{FiberBundle2020, rowlandFiberBundle}. A fiber bundle is a structure $(E, B, \pi, F)$  consisting of topological total space $E$, base space $K$, fiber space $F$ and the map from total space to base space:

@@

where there is a bijection from $F$ to every fiber $F_b$ over point $b \in B$ in $E$ and the function $\pi: E \rightarrow B$ is the map into the $B$ quotient space of $E$. By defintion of a fiber bundle, $\pi$ is always a mapping from total space to base space, independent of the points $p \in E$, and therefore we call this mapping $\pi$ for all the fiber bundles in the model.

\subsubsection{Base space}
\begin{figure}[H]
    \label{fig:base_space}
    \includegraphics[width=0.4\linewidth]{figures/sections/math/k_qspace.png}
    %%include second figure where fiber goes otherway
\end{figure}

%%% add in motivation - encodes the connectiveity of the points, choice of split is intrinsinc to how we want to view the data

$B$ is the quotient space of $E$, meaning it is the set of equivalence classes of elements $p$ in $E$ defined via the map $\pi: E \rightarrow B$ that sends each $p \in E$ to its equivalence class in $[p] \in B$ \cite{QuotientSpaceTopology2020,QuotientSpaceTopology2020}.

As shown in figure~\ref{fig:kquote}, the fibers $F$ divide $E$ into smaller spaces consisisting of $F$ and an open set neighborhood \cite{FiberBundle2020, rowlandFiberBundle}  around $F$.
%%so why topology? 
\paragraph{Example}
%%% pull up data structures like discrete, continous, etc w/ base space 
%%% little drawing dots, line, plane, cube

\begin{figure}[ht]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_1k.png}
    %% add box around neighboring P and Map
    \label{fig:k_data}
\end{figure}

\begin{figure}[H]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_2k.png}
\end{figure}

@@

\subsubsection{Fiber}
%% pull fiber stuff up here? what is it? why? fiber as cartesian product of sets
%% actual values are in the fiber 
%% gonna view spaces as simple piece put together 
%% Our fibers are cross products of (column name, domain) as described by Spivak's \cite{spivakSIMPLICIALDATABASES}.

@@

\subsubsection{Section}
%% add context
The section $f$ is the mapping from base space to total space $f: B\rightarrow E$ 
\begin{equation}
    \label{eq:section}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & B \arrow[u, "f"', bend right]
    \end{tikzcd}
\end{equation}

@@

In a locally trivial fiber bundle, $E = B \times F$ \cite{rowlandFiberBundle,FiberBundle2020}:
\begin{equation}
    \label{eq:section_return}
f(b) = (b, g(b))
\end{equation}

@@

\subsubsection{Sheaf and Stalk}
%% add motivation - zoom, subsection, streaming, also for realizing visualization, restricting to that piece - need to know near point w/o 
As described in equation~\ref{eq:local_trivial}, there is a local space $U \subset B$ around every $b$. The inclusion map $\iota: U \rightarrow B$ can be pulled back such that $\iota^{*}E$ is the space of $E$ restricted over $U$. 
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*E \arrow[d, "\pi"']           & E \arrow[d, "\pi"'] \arrow[l, "\iota^*"']         \\
        U \arrow[u, "\iota^*f"', bend right] & B \arrow[u, "f"', bend right] \arrow[l, "\iota"']
        \end{tikzcd}
\end{equation}

The localized section of fibers $\iota^*f: U \rightarrow \iota^*E$ is the sheaf $\mathcal{O}(E)$ with germ of $\xi^*f$ which is the localized section. The neighborhood of points the sheaf lies over is the stalk $\mathcal{F}_b$ \cite{StalkSheaf2019,spanier1989algebraic} which is a ...$\dots$
. Restricting the artist to the sheaf means the artist knows the data in $F$ and also has access to derivatives of the data. This property is useful for some visual transformations. 
%% gives all direvatives but we need the jet which is Add stuff about jets

\subsubsection{Triangulization}
%%drop down to after the fiber section
The base space $K$ is a representation of the connectivity of the data, specifically whether the points in $E$ are discrete or sampled from a continous space. The same dataset can be expressed with different $K$.

%% k is a triangularizable topological space 
One way to represent the topological space K is as a set composed of simplices, such as those shown in figure~\ref{fig:simplex}. Simplices are a way of encoding the connectivity of each observation ($\sigma(k)$) to another: %%formula for simplacial complexes/simplacial sets

\subsection{Prerender Space}
\label{sec:display}

We model the prerender space on which lives on ideal version of the visulaization as a fiber bundle (H, S, $\pi$, D). $H$ is the predisplay space, with a fiber $D$ dependent on the target physical display and a base space of $S$.

\subsubsection{Base space}
%%pull up to the very top that dim(k) < screen space, need to thicken therefore S
$K$ can be considered a subspace of the screen base space $S$ such that $\xi: S \rightarrow K$ is a deformation retraction \cite{RetractionTopology2020}
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] & H \arrow[d, "\pi"'] \\
        K                   & S \arrow[l, "\xi"']
        \end{tikzcd}
\end{equation}

@@

\begin{equation}
    \label{eq:artist}
    \begin{tikzcd}
        E \arrow[r, "\nu"] \arrow[rd, "\pi"'] & V \arrow[d, "\pi"] & \xi^*V \arrow[r, "Q"] \arrow[d, "\xi^*\pi"'] \arrow[l, "\xi^*"'] & H \arrow[ld, "\pi"] \\
                                              & K                  & S \arrow[l, "\xi"']                                              &                    
        \end{tikzcd}
\end{equation}
%%map backwords via pullback can copy values over k over and over to corresponding S_k

@@

\subsubsection{Assembling Marks}
\begin{figure}[H]
    \includegraphics[width=0.4\linewidth]{figures/sections/math/path_of_q}
    \label{fig:q}
    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$.}  
\end{figure}

The assembly function $Q: \Gamma(V) \rightarrow \Gamma(H)$ composites the visual variables in $V$ into an element in $H$. Given a monoid action on a set in $V$, there should be a monoid action on the corresponding $Q(\Gamma(V))$. While $\Gamma(V)$ holds for all cases, we can specialize to the bundle $V$ or the sheaf $\mathcal(V)$ depending on the specific visualization.

@@

For each region $s$ in the display space $H$, the mark  \cite{bertinIIPropertiesGraphic2011,carpendaleVisualRepresentationSemiology} it belongs to can be found by mapping $s$ back to $K$ via the lookup on $S$ then taking $\xi(s)$ back to a point on $k \in K$ which lies on the path connected component $J \subset K$.

%% defined J_{k} first, then pull down to here, then can do \rho S_j
\begin{equation}
    \begin{tikzcd}
        H \arrow[r, shift left] & S \arrow[l, "\rho(\xi^{-1}(J))", shift left] \arrow[rr, "\xi(s)", shift left] &  & J_{k} =  \{j \in K \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=k \text{ and }\gamma(1)=j\} \arrow[ll, "\xi^{-1}(J)", shift left]
    \end{tikzcd}
    \label{eq:mark}
\end{equation}
where the path\cite{ConnectedSpace2020}  $\gamma$ from $k$ to $j$ is a continous function from the interval [0,1]. To get back to the display space $H$ from $J$, the inverse image of $J \in S, \xi^{-1}(J)$ is pushed back to $S$, and then $\rho(\xi^{-1}(J))$ maps it into $R^{7}$ such that $Q\circ\xi^*\mu(\xi^{-1}(J))$ generates a graphical mark.

>>>>

%% convention is base space \times fiber bundle
%% change B to K
\section{Notation \& Definitions}

%% give the why of the math after every mention of why

In this section we introduce a mathematical description of the visualization pipeline where artist $\mathcal{A}$ functions transform data  $\mathcal{E}$ to an intermediate representation in a prerendered display space $\mathcal{H}$.

\begin{equation}
    \label{eq:artist}
    \mathcal{A}: \mathcal{E} \rightarrow \mathcal{H}
\end{equation}

We use fiber bundles\cite{FiberBundle2020, rowlandFiberBundle} to model data and graphics because they are products of a topological total space that encodes points and a topological base space that can encode the connectivity of those points:
\begin{itemize}
\item $E$ is a locally trivial fiber bundle over $K$ representing data space.
\item $H$ is a fiber bundle over $S$ representing visual space
\item $K$ and $S$ are a triangulizable topological space or a CW complex encoding the connectivity of points in $E$ and $H$ respectively
\end{itemize}

The fiber bundles mentioned in this work are assumed to be locally trivial\cite{spanier1989algebraic,LocallyTrivialFibre}.

\subsection{Data Fiber Bundles}
As proposed by Butler \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, we model data as a fiber bundle ($E$, $K$, $\pi$, $F$)

@@

with topological total space $E$, base space $K$, fiber space $F$, and the map from total space to base space $\pi: E \rightarrow K$. Maps from $K$ to $E$ are called sections and select specific points in $K$. The global space of sections in $E$ is $\Gamma(E)$.

\paragraph{example}

\begin{figure}[H]
    \label{fig:data_fiber_bundle}
    \includegraphics[width=1\linewidth]{figures/sections/math/fiberbundle.png}
    \caption{$F$ is an $n\times m$ space of points. The section $\tau_1$ returns the blue points,  while $\tau_2$ returns the purple points. $\Gamma(E)$ is the set of all sections, including $\tau_1$ and $\tau_2$}.  
\end{figure}

The fibers $F$ in figure~\ref{fig:data_fiber_bundle} lie in some total space $E$ and are points that lie in a 2D space. For example, this space can be $\mathrm{temperature}\times \mathrm{pressure}$ or $\mathrm{time} \times \mathrm{precipitation}$. $K$ is a line on the interval [0,1] and indicates that the points lie on a 1D continous space, but we derive whether the data is a distribution or a timeseries or a function from the fields in $F$. The sections $\tau_1$ and $\tau_2$ are distinct maps from $k$ to the blue and purple points in $E$ respectively.

\subsubsection{Base Space}
\label{sec:base_data}
\begin{figure}[H]
    %%\includegraphics[\textwidth]{}
    \label{fig:base_space_types}
    %%% pull up data structures like discrete, continous, etc w/ base space 
    %%% little drawing dots, line, plane, cube
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

Data can be discrete observations, timeseries, maps, fields \cite{munznerWhatDataAbstraction2014} and $K$ is a set of points $k$ that can act as keys from a representative space, such as seen in figure~\ref{fig:base_space_types}, to the data values in $E$ \cite{munznerWhatDataAbstraction2014}.

\begin{figure}[H]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/sections/math/k_qspace.png}
    %%include second figure where fiber goes otherway
\end{figure}

$K$ and $F$ are not intrinsic to $E$, rather they are choices in how $E$ is subdivided\cite{QuotientSpaceTopology2020,QuotientSpaceTopology2020}. In figure~\ref{fig:base_space_div} we can divide a rectangular base space such that there is a short fiber and long base space or a long fiber and short base space. This is analogous to long and wide forms of the same table \cite{wickham2014tidy}.

\paragraph{Example}
\begin{figure}[ht]
    \includegraphics[width=\textwidth]{figures/sections/math/temp_1k.png}
    %% add box around neighboring P and Map
    \label{fig:k_data}
\end{figure}

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/sections/math/temp_2k.png}
\end{figure}

@@

\subsubsection{Triangulization}
\label{sec:triangulization}

The base space $K$ is a representation of the connectivity of the data, specifically whether the points in $E$ are discrete or sampled from a continous space. The same dataset can be expressed with different $K$.

K is a triangularizable topological space; one triangularization scheme is as a set composed of simplices, such as those shown in figure~\ref{fig:simplex}.

\subsubsection{Fiber Space}
\label{sec:fiber_data}
%% pull fiber stuff up here? what is it? why? fiber as cartesian product of sets
%% actual values are in the fiber 
%% gonna view spaces as simple piece put together 
%% Our fibers are cross products of (column name, domain) as described by Spivak's \cite{spivakSIMPLICIALDATABASES}.

@@

\subsubsection{Section}
\label{sec:section_data}
%% add context
The section $\tau$ is the mapping from base space to total space $\tau: K\rightarrow E$ 
\begin{equation}
    \label{eq:section}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & B \arrow[u, "f"', bend right]
    \end{tikzcd}
\end{equation}

@@

In a trivial fiber bundle, $E = K \times F$ \cite{rowlandFiberBundle,FiberBundle2020}:
\begin{equation}
    \label{eq:section_return}
f(b) = (b, g(b))
\end{equation}

@@

\subsubsection{Sheaf and Stalk}
%% add motivation - zoom, subsection, streaming, also for realizing visualization, restricting to that piece - need to know near point w/o

As described in equation~\ref{eq:local_trivial}, there is a local space $U \subset B$ around every $b$. The inclusion map $\iota: U \rightarrow B$ can be pulled back such that $\iota^{*}E$ is the space of $E$ restricted over $U$. 
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*E \arrow[d, "\pi"']           & E \arrow[d, "\pi"'] \arrow[l, "\iota^*"']         \\
        U \arrow[u, "\iota^*f"', bend right] & B \arrow[u, "f"', bend right] \arrow[l, "\iota"']
        \end{tikzcd}
\end{equation}

The localized section of fibers $\iota^*f: U \rightarrow \iota^*E$ is the sheaf $\mathcal{O}(E)$ with germ of $\xi^*f$ which is the localized section. The neighborhood of points the sheaf lies over is the stalk $\mathcal{F}_b$ \cite{StalkSheaf2019,spanier1989algebraic} which is a ...$\dots$
. Restricting the artist to the sheaf means the artist knows the data in $F$ and also has access to derivatives of the data. This property is useful for some visual transformations. 
%% gives all direvatives but we need the jet which is add stuff about jets

\subsection{Prerender Space}
\label{sec:display}
We define a graphic space $H$ such that we do not have to assume the physical output space of the renderer. This means that the graphic in $H$ can be output to a screen or 3D printed space or a dome.

We model the prerender space as a fiber bundle (H, S, $\pi$, D). $H$ is the predisplay space, with a fiber $D$ dependent on the target display and a base space of $S$.

\subsubsection{Base space}
The underlying topology $S$ of a graphic often needs more dimensions than the data topology $K$ because of the specifications of the display space. For example, a line plot on a plane (such as a screen or a piece of paper) by necessity needs to also have a thickness so that it is visible, which maps back to a set of connected points in $H$. The topology of these connected points is therefore the region $s \subset S$ such that $\xi: S \rightarrow K$ is a deformation retraction \cite{RetractionTopology2020}
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] & H \arrow[d, "\pi"'] \\
        K                   & S \arrow[l, "\xi"']
        \end{tikzcd}
\end{equation}

@@

\begin{equation}
    \label{eq:artist}
    \begin{tikzcd}
        E \arrow[r, "\nu"] \arrow[rd, "\pi"'] & V \arrow[d, "\pi"] & \xi^*V \arrow[r, "Q"] \arrow[d, "\xi^*\pi"'] \arrow[l, "\xi^*"'] & H \arrow[ld, "\pi"] \\
                                              & K                  & S \arrow[l, "\xi"']                                              &                    
        \end{tikzcd}
\end{equation}
The pullback map $\xi^*$ copies each value in $V$ over $k$ to $s$ in corresponding $S_k$ such that $\xi^*V$ can have multiple values that map to one value in $V$.

@@

\subsubsection{Assembling Marks}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/sections/math/path_of_q}
    \label{fig:artist_q}
    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$.}  
\end{figure}

As shown in figure~\ref{fig:artist_q}, $Q$ takes the individual fields in $V$ as input and outputs a single piece of a graphic on $H$. As with $\nu$, the constraint on $Q$ is that for every monoid actions on the input in $V$ there is a corresponding monoid action on the output in $H$.

@@

To output a mark  \cite{bertinIIPropertiesGraphic2011,carpendaleVisualRepresentationSemiology}, $Q$ is called with all the regions $s$ that map back to a set of connected components $J \subset K$:
\begin{equation}
J = \{j \in K \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=k \text{ and }\gamma(1)=j\}
\end{equation}
where the path\cite{ConnectedSpace2020}  $\gamma$ from $k$ to $j$ is a continous function from the interval [0,1].

We define the mark as the graphic generated by $Q(S_j)$:

\begin{equation}
    \begin{tikzcd}
        H \arrow[r, shift left] & S_j \arrow[rr, "\xi(s)", shift left] \arrow[l, "\rho(S_j)"] &  & J_{k} \arrow[ll, "\xi^{-1}(J)"]
        \end{tikzcd}
    \label{eq:mark}
\end{equation}

where the set $j \subset J$ is the set of marks in the graphic.
